The CMS Data Aggregation System

### 2010-05-04 doc/das_paper/paper.tex
Relocate docs for das paper into separate folder
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@851 4525493e-7705-40b1-a816-d608a930855b

Once such a conglomerate of data-services start operating the obvious
question arises: how do we find the desired information across multiple data-services
in our distributed environment? Even though the individual data-services were designed
to answer specific questions about the data they serve, the ability to search and relate
information among them was remains a tedious human task. The growing amount of information
and desire to make cross-service queries led us to design and develop the new
Data Aggregation System.

@@

\subsection{DAS core\label{DAS_core}}
The DAS core library is responsible for data retrieval on demand. 
It has the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server was done via the REST \cite{REST} interface. 
Each input query was recorded into the Analytics DB for further 
analysis, and was decomposed into a set of conditions and/or selection keys.
First the results were looked up in a cache and delivered to the requester
if superset of input conditions was found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, was constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and stored results into the cache. At this step we also
performed necessary data aggregation for records with similar key-value
pairs. The data-service plug-in implementations are similar and their 
logic is largely abstracted, apart from a few extensions such as specific
security modules for data-services accessible via private network.
Finally, the results were retrieved from the cache and delivered to the user.

@@

\begin{table*}[hbt]
\centering
\begin{tabular}{lllll}\hline
\hline
System & Format & Records & Elapsed time & Elapsed time \\
& & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. The PhEDEx
system shows fetched (186145) and merged (180439) records. 
The total number of DAS records are calculated as number of 
aggregated records plus left overs, the non-matched records from both systems. 
The look-up time shown in last column represents query look-up 
time (2.05 sec) and time required to get all records written 
to disk (50.7 sec).}
\label{DAS_benchmark}
\end{table*}

@@

The results of the stress tests have shown that we can achieve a throughput of
$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
$\sim$7600 docs/sec for reading operations in DAS, which is suitable for our 
needs. Please note, we have relatively small audience of users
who perform complex queries, while search engines designed for
a large number of users asking for imprecise, simple data.
Since all tests were performed on a single CPU, further improvements from
expanding to multiple CPUs are expected. In addition, more
performance advantage can be gained by writing dedicated C/C++
python extensions for data intensive operations.

>>>>

Once such a conglomerate of data-services start operating the obvious
question arises: how do we find the desired information across multiple data-services
in our distributed environment? Even though the individual data-services were designed
to answer specific questions about the data they serve, the ability to search and relate
information among them remains a tedious human task. The growing amount of information
and desire to make cross-service queries led us to design and develop the new
Data Aggregation System.

@@

\subsection{DAS core\label{DAS_core}}
The DAS core library is responsible for data retrieval on demand. 
It has the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server is done via the REST \cite{REST} interface. 
Each input query is recorded into the Analytics DB for further 
analysis, and is decomposed into a set of conditions and/or selection keys.
First the results are looked up in a cache and delivered to the requester
if superset of input conditions is found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, is constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and stored results into the cache. At this step we also
performed necessary data aggregation for records with similar key-value
pairs. The data-service plug-in implementations are similar and their 
logic is largely abstracted, apart from a few extensions such as specific
security modules for data-services accessible via private network.
Finally, the results are retrieved from the cache and delivered to the user.

@@

\begin{table*}[hbt]
\centering
\begin{tabular}{lllll}\hline
\hline
System & Format & Records & Elapsed time & Elapsed time \\
& & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
DAS (total) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. The PhEDEx
system shows fetched (186145) and merged (180439) records. 
The last row shows aggregated results and total time
spent in all components (DBS+PhEDEx+aggregation).
The total number of DAS records are calculated as number of 
aggregated records plus left overs, the non-matched records from both systems. 
The look-up time shown in last column represents query look-up 
time (2.05 sec) and time required to get all records written 
to disk (50.7 sec).}
\label{DAS_benchmark}
\end{table*}

@@

The results of the stress tests have shown that we can achieve a throughput of
$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
$\sim$7600 docs/sec for reading and writing DAS records to disk,
which is suitable for our 
needs. Please note, we have relatively small audience of users
who perform complex queries and get precise results, while search 
engines are designed for a large number of users asking for imprecise, simple data.
Since all tests were performed on a single CPU, further improvements from
expanding to multiple CPUs are expected. In addition, more
performance advantage can be gained by writing dedicated C/C++
python extensions for data intensive operations.
